---
title: "STAT 345 Homework 1"
output:
  html_document: default
  pdf_document: default
---
Homework 1 
By: Eli Temanson

1. The data set at [http://www.stats.uwo.ca/faculty/braun/data/rnf6080.dat] records hourly rainfall at a certain location in Canada, every day from 1960 to 1980.
    a. First, we need to load the data set into R using the command `read.table()`. Use the help function to learn what arguments this function takes. Once you have the necessary input, load the data set into R and make it a data frame called `rain.df`.
    b. How many rows and columns does `rain.df` have? (If there are not 5070 rows and 27 columns, something is wrong; check the previous part to see what might have gone wrong in the previous part.)
```{r}
rain.df <- read.table("http://www.stats.uwo.ca/faculty/braun/data/rnf6080.dat")
```
    
  c. What are the names of the columns of `rain.df`?
  **Name of Columns are V1 V1 ... V27 and rows are 1 through 5070**
```{r}
head(rain.df)
```

  d. What is the value of row 5, column 7 of `rain.df`?
  **Value of row 5, column 7 is 0.**
```{r}
rain.df[5,7]
```

  e. Display the second row of `rain.df` in its entirety.
```{r}
rain.df[2,]
```

  f. Explain what this command does:
```
names(rain.df) <- c("year","month","day",seq(0,23))
```
  by running it on your data and examining the object. (You may find the display functions `head()` and `tail()` useful here.) Is it clear now what the last 24 columns represent?

  **The second command below changes the first three columns to "year", "month", "day" and the rest of the columns to order from (0,23) instead of (0,27). The last 24 columns represent the hours in a day.**
```{r}
names(rain.df) <- c("year","month","day",seq(0,23))
head(rain.df)
```

  g. Create a new column in the data frame called `daily`, which is the sum of the rightmost 24 columns. With this column, create a histogram of the values in this column, which are supposed to be daily rainfall values. What is wrong with this picture?  
```{r}
rain.df$daily <- rowSums(rain.df[,4:27])
head(rain.df)
hist(rain.df$daily)
```

  h. Create a new data frame `rain.df.fixed` that takes the original and fixes it for the apparent flaw you have discovered. Having done this, produce a new histogram with the corrected data and explain why this is more reasonable.
```{r}
rain.df.fixed <- rain.df
for(i in 1:length(rain.df$daily))
  {
  if(rain.df$daily[i] < 0)
  {
    rain.df.fixed$daily[i] <- (rain.df$daily[i] = 0)
  }
}
hist(rain.df.fixed$daily[!rain.df.fixed$daily==0],breaks=75)

```

2. Syntax and class-typing.
    a. For each of the following commands, either explain why they should be errors, or explain the non-erroneous result. 
    **For the second command below the max() function on a vector of strings doesn't make sense, "7" is the output which isn't the numeric max() but the largest of the first place holder.**
    **The third command sort() does something interesting, it sorts the strings by the largest first place holder instead of true numeric value.**
    **Sum() function sends an error becuase you cannot sum characters, you would need to use something like cbind()**
```
vector1 <- c("5", "12", "7", "32")
max(vector1)
sort(vector1)
sum(vector1)
```
    b. For the next series of commands, either explain their results, or why they should produce errors.
    **The second command will not work due to the non-numeric components of the vector. Since the vector contains a character "5" it made the entire vector characters. **
    **The fourth command works well because of the difference in vectors and dataframes. Dataframes hold on to and seperate different types of variables.**
    **Lists look similar to the vector format in which we don't have to specify row or column just location in the list[i]. We can also cast from string to numeric by using double brackets [[i]], this is why the last command below outputs an error.** 
    
```
vector2 <- c("5",7,12)
vector2[2] + vector2[3]

dataframe3 <- data.frame(z1="5",z2=7,z3=12)
dataframe3[1,2] + dataframe3[1,3]

list4 <- list(z1="6", z2=42, z3="49", z4=126)
list4[[2]]+list4[[4]]
list4[2]+list4[4]
```

3. Working with functions and operators.
    a. The colon operator will create a sequence of integers in order. It is a special case of the function `seq()` which you saw earlier in this assignment. Using the help command `?seq` to learn about the function, design an expression that will give you the sequence of numbers from 1 to 10000 in increments of 372. Design another that will give you a sequence between 1 and 10000 that is exactly 50 numbers in length.
    ```{r}
    seq1 <- seq(from=1,to=10000,by=372)
    seq2 <- seq(from=1,to=10000,length.out=50)
    ```
    b. The function `rep()` repeats a vector some number of times. Explain the difference between `rep(1:3, times=3)` and `rep(1:3, each=3)`.
    **The difference between the two arises from the use of 'times=3' vs. 'each=3'. As you see below 'times=3' cycles through the 1:3 three times, where the 'each=3' outputs each increment three times. **
```{r}
rep(1:3, times=3)
rep(1:3, each=3)
```